---
title: "RKDB README"
author: 
output:
  md_document:
    variant: gfm
params:
  server: localhost
  port: 5000
---

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "figures/"
)
```

## R client for kdb+

Execute kdb+ queries from R for advanced high-performance analytics.

A KX [Fusion interface](https://code.kx.com/q/interfaces#fusion)

```{r hooks, include=FALSE}
# add stringr
if(! 'gsubfn' %in% rownames(installed.packages())) install.packages('gsubfn')
library(gsubfn)
default_source_hook <- knitr::knit_hooks$get('source')
knitr::knit_hooks$set(source = function(x, options) {
  x <- gsubfn(x = x, pattern = "params\\$`?([\\w_]+)`?", function(y) {
    y <- get(y, params)
    ifelse(is.character(y), paste0("'", y, "'"), y)
  })
  default_source_hook(x, options)
})
```

# Requirements
R 3.3.x and later

# Installation
```{r install 1, echo=TRUE, message=FALSE, warning=FALSE}
# remove old package
if('qserver' %in% rownames(installed.packages())) remove.packages('qserver')
# install devtools
if(! 'devtools' %in% rownames(installed.packages())) install.packages('devtools')
library(devtools)
# install rkdb
devtools::install_github('kxsystems/rkdb', quiet=TRUE)
# to install rkdb of particular release
# devtools::install_github('kxsystems/rkdb@0.10.0', quiet=TRUE)
library(rkdb)
```

# First steps
## Setting up a connection
Start kdb+ process to test the installation
```{r start kdb, echo=TRUE, eval=FALSE}
q -p params$port
```

Open a connection to it
```{r open connection, echo=TRUE}
h <- open_connection(params$server,params$port)
```

## Hello kdb

You can evaluate any kdb expression and its result will come back to R:
```{r abc 1, echo=TRUE}
execute(h, '1+1')
```

Assigning a variable in q workspace also works:
```{r abc 2, echo=TRUE}
execute(h, 'x:1+1') #assign x to 2
execute(h, 'x') # get back the value
```

# Getting data from kdb to R

kdb+ uses some basic types that might not have a direct equivalent in R. Note also that this is not a bijective operation. The conversions (from kdb to R) are: 

```{r atoms 0, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
cat("
| kdb/q                       | r            |
|-----------------------------|--------------|
| boolean                     | logical      |
| byte                        | raw          |
| short                       | integer      |
| int                         | integer      |
| long                        | integer64    |
| real                        | numeric      |
| float                       | numeric      |
| char                        | character    |
| symbol                      | character    |
| timestamp                   | nanotime     |
| month                       | integer      |
| date                        | Date         |
| datetime                    | POSIXct      |
| timespan                    | integer64    |
| minute                      | difftime     |
| second                      | difftime     |
| time                        | integer      |
| enumeration                 | character    |
| table                       | data.frame   |
| keyed table                 | data.frame   |
| dictionary (mixed types)    | named list   |
| dictionary (same types)     | named vector |
| function                    | character    |
| list (same types)           | vector       | 
| list (same 'complex' types) | list         | 
| list (different types)      | list         | 
")
```

## Computing on kdb

rkdb provides a convenient way to retrieve computation done on the kdb side so you can have the best of both worlds:
```{r kdb side example 1, echo=TRUE, warning=FALSE, message=FALSE}
kdb <- '
t: ([] x:1000#`a`b`c;y:1000#1f*til 10;z:1000#1f*til 4);
select sum y, dev z by x from t
'

execute(h, kdb)
```

One can for instance use R graphical capabilities:
```{r example, echo=TRUE, warning=FALSE, message=FALSE}
kdb <- '
t: ([] x:1000#`a`b`c;y:1000#1f*til 10;z:1000#1f*til 4);
select y,z from t where x=`a
'

DF <- execute(h, kdb)
plot(DF$y, DF$z, main='scatter plot', xlab='y values', ylab='z values')
```

# Getting data from R to kdb
## Evaluating kdb expressions using R objects

You can call kdb functions with R objects as arguments, those will be passed and converted to native kdb+ data types, and the kdb expression will be evaluated:
```{r use r obj 1, echo=TRUE}
execute(h, "raze", list(c(1,2,3), c(4,5,6)))
execute(h, "+", 2, 5)
execute(h,'{`tmp set x}',data.frame(a=c(1,2,3),b=c("a","b","b")))
```

As an example here is how you can use the left-join function on 2 data.frames:
```{r ex data.frame, echo=TRUE, warning=FALSE, message=FALSE}
DF1 <- data.frame(x=c('x','x','y','y'), y=1:4)
DF2 <- data.frame(x=c('x','y','z'), z=seq(10,30,10))
execute(h, "{[x;y] x lj `x xkey y}", DF1, DF2)
```

## :open_file_folder: Documentation

*   [User guide](docs/README.md)
*   [Using R with kdb+](r-and-q.md)
*   [Other methods](other-methods.md)

---
See [code.kx.com/q/interfaces](https://code.kx.com/q/interfaces) for other interfaces to and from kdb+.
